---
title: "Class Activity 13"
author: "Arthur Viegas Eguia"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      size = "small", 
                      collapse = TRUE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      error = TRUE)

# load the necessary libraries
library(wordcloud)
library(reshape2)
library(tidyverse)
library(tidyr)
library(tidytext)
library(dplyr)
```


\vspace*{1in}


## Group Activity 1

### a. Variance and Skewness 

The variance of a random variable $x$ is defined as: 

$$\operatorname{Var}(x)=\frac{1}{n-1} \sum_{i=1}^{n}\left(x_{i}-\bar{x}\right)^{2}$$

where $x_i = (\sum_i^n x_i)/n$ is the sample mean. Also, the skewness of the random variable $x$ is defined as:

$$\operatorname{Skew}(x)=\frac{\frac{1}{n-2}\left(\sum_{i=1}^{n}\left(x_{i}-\bar{x}\right)^{3}\right)}{\operatorname{Var}(x)^{3 /2}}$$

Please write functions to calculate the variance and skewness of $\{12, 45, 54, 34, 56, 30, 67, \text{NA}\}$.

```{r}
x <- c(12, 45, 54, 34, 56, 30, 67, NA)
```




```{r}
# function to calculate the variance of a vector
var <- function(x){
  x <- na.omit(x)
  res <- (sum((x - mean(x))^2))/(length(x) - 1)
  return(res)
}
```


```{r}
var(x)
```



```{r}
# function to calculate the skewness of a vector
skewness <- function(x){
  x <- na.omit(x)
  res <- sum((x - mean(x))^3)/((length(x) - 2) * var(x)^(3/2))
  return(res)
}
```


```{r}
skewness(x)
```


--------------------------------------------------------------------

## Group Activity 2


```{r}
musical_instr_reviews <-  read_csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/musicreviews.csv") %>% 
  rename(ratingOverall = overall)
glimpse(musical_instr_reviews)
```

### a. Write a function to filter the dataset based on the provided rating:

```{r}
filter_reviews_by_rating <- function(dataset, review){
  dataset %>% filter(ratingOverall == review)
}
filtered_data <- filter_reviews_by_rating(musical_instr_reviews, 4)
```


### b. Write a function to process the text and remove stop words:

```{r}
process_text <- function(data){
  data %>%
    select(reviewText) %>%
    unnest_tokens(output = word,
                  input = reviewText) %>% 
    anti_join(stop_words)
}
process_text(filtered_data) -> process_data_4
```


### c. Write a function to join the processed text with sentiment data and create a word count table.

```{r}
create_word_count_table <- function(data){
  data %>% 
    inner_join(get_sentiments("bing")) %>% 
  count(word, sentiment, sort = TRUE) %>% 
  reshape2::acast(word ~ sentiment, value.var = "n", fill = 0)
}


```


### d. Create the final function that takes the rating and number of words as input arguments and returns a word cloud plot. 

```{r}
library(wordcloud)
library(reshape2)

word_cloud <- function(rating, num.words) {
 rating = as.numeric(rating)
 num.words = as.numeric(num.words)
 if(rating >= 1 & rating <= 5){
   filtered_data <- filter_reviews_by_rating(musical_instr_reviews, rating)
   processed_data <- process_text(filtered_data)
   word_table <- create_word_count_table(processed_data)
   wordcloud::comparison.cloud(word_table, colors = c("blue", "green"), 
                               scale = c(2, 0.5),
                               max.words = num.words,
                               title.size = 2)
 }
 else warning("Please enter a number from 1 to 5")
}
word_cloud(rating = "4", num.words = 50)
```







