---
title: "Class Activity 9"
author: "Arthur Viegas Eguia"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      size = "small", 
                      collapse = TRUE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      error = TRUE)

# load the necessary libraries
library(tidyverse)
library(janitor)
library(stringr)
library(tidyr)
library(readr)
```


\vspace*{1in}

## Problem 1

a) Use `read_csv()` to import the `desserts` data set from [GitHub url](https://raw.githubusercontent.com/deepbas/statdatasets/main/desserts.csv). Use `glimpse` to see if the data import is alright.

```{r}
url <- "https://raw.githubusercontent.com/deepbas/statdatasets/main/desserts.csv"
desserts <- readr::read_csv(url)
glimpse(desserts)
```

Does everything look good? Import the dataset with correct data types, if needed. Fix the problems, if any.

```{r}
# your r-code
desserts1 <- read_csv(url, col_types = list(
    technical=col_number(),
    uk_airdate = col_date()
  ))
glimpse(desserts1)
problems(desserts1)
```
```{r}
desserts2 <- read_csv(url, col_types = list(
  technical = col_number(),
  uk_airdate = col_date(format = "%d %B %Y")
))
problems(desserts2)
```

```{r}
desserts3 <- read_csv(url, col_types = list(
  technical = col_number(),
  uk_airdate = col_date(format = "%d %B %Y")
),
na = c("N/A", "", "unknown", "NA")
)
problems(desserts3)
```



## Problem 2

a) Use the appropriate `read_<type>()` function to import the datasets provided in the [course webpage](https://stat220-spring24.netlify.app/datasets). Identify and fix any issues you encounter.


If you hit any errors/problems, be sure to explore them and identify the issue, even if you can't "fix" it.


```{r}
# your code for simple 1
simple <- readr::read_csv("data/simple-1.dat")
#write_csv(simple, "data/simple_copy.csv")
```


```{r}
# your code for mild 1
read_csv("data/mild-1.csv")
mild1 <- read_delim("data/mild-1.csv")
```


```{r}
# your code for tricky 1
tricky1 <- read_csv("data/tricky-1.csv")
problems(tricky1)

```


The issue is that we have missing values that aren't specifically included in the rows 4 and 7 of the **original** file (so rows 3 and 6 once we load the data). We can fix this with post processing

```{r}
# your post-processing code
tricky1[3, ] <- c(tricky1[3, 1:2], NA, tricky1[3, 3:4])
tricky1[6, ] <- c(tricky1[6, 1], NA, tricky1[6, 2:4])

tricky1
```


```{r}
# your code for tricky 2
tricky2 <- read_csv("data/tricky-2.csv")
problems(tricky2)
tricky2
```

Why are there state abbreviations in the latitude column?

```{r, eval=FALSE}
# your post-processing code
tricky2_part1 <- read_csv("data/tricky-2.csv", n_max = 7)



tricky2_part1 <- tricky2_part1 %>% separate(city, c("city, state"), sep=", ") 
names <- colnames(tricky2_part1)
tricky2_part1

tricky2_part2 <- read_csv("data/tricky-2.csv", skip = 8, col_names = names)
full_join(tricky2_part1, tricky2_part2) -> fullCleanData

```



## Acknowledgemant

Parts of the activities are adapted from similar activity written by Adam Loy.



