---
title: "Class Activity 15"
author: "Arthur Viegas Eguia"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      size = "small", 
                      collapse = TRUE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      error = TRUE)

# load the necessary libraries
library(tidyverse)
library(stringr)
library(purrr)


energy <- read_csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/energy.csv",
                    col_type = cols(
                     .default = col_double(), 
                      Timestamp = col_datetime(format = ""),
                      dayWeek = col_factor(levels=c("Mon","Tues","Wed","Thurs","Fri","Sat","Sun"))
                     ))
```


\vspace*{1in}


## Group Activity 1

### a. Write a for loop to iterate over the columns of the 'energy' dataset and print the names of all columns containing the string "House". Please use the function `colnames()` to extract the column names and store the results in a list.

*Answer:*

```{r}
results <- list()
for(i in colnames(energy)){
  if(str_detect(i, "House")){
    results[[length(results) + 1]] <- i 
  }  
}

results %>% unlist() -> houses
```





### b. Using a for loop, calculate and print the mean energy consumption of houses you identified in part a.

```{r}
mean_energy <- c()
for(i in houses){
  mean_val <- mean(energy[[i]], na.rm = TRUE)
  mean_energy <- c(mean_energy, mean_val)
}
mean_energy
```


-------------------------------------------------------------------------------------

## Group Activity 2


**1.** Make a data frame of quantiles for `energy` buildings in columns 9-90 (you will need `na.rm = TRUE`)

```{r}
qdf1 <- energy %>%
  select(9:90) %>% 
  map_dfc(quantile,
          probs = seq(0.1, 0.9, 0.1),
          na.rm =TRUE) 
qdf1
```




**2.**  Add a variable to identify the quantile


```{r}
qdf1 %>%
  mutate(stat = str_c("quantile_",
                      seq(10, 90, 10)),
         .before=1) ->qdf2
qdf2
```







**3.**  Reshape the data frame to make variables `stat` (describing the quantile), `building` and `quant` (quantile value)




```{r}
qdf2 %>%
  qdf3 <- pivot_longer(names_to = "Building",
                      values_to="Quantiles",
                      cols = -1)
qdf3
```





**4.**  Plot the KWH value for each quantile on the x-axis for the buildings `Sayles-Hill`, `Center_for_Mathematics_&_Computing`, `Olin_Hall_of_Science`



```{r}
qdf3 %>% 
  filter(Building %in% c("Sayles-Hill",
                         "Center_for_Mathematics_&_Computing",
                         "Olin_Hall_of_Science")) %>% 
  ggplot(aes(x = Quantiles, y = readr::parse_number(stat), color = Building)) + 
  geom_point() +
  geom_line(aes(color=Building)) +
  scale_y_continuous(breaks = seq(10, 90, 10)) +
  labs(y = "Percentiles (%)",
       x = "Energy KWH")

```








