---
title: "Homework 5"
output: pdf_document
---

## Name: Arthur Viegas Eguia

## I worked with: Sarah, Allison, and Jenna at the stats help lab

**Click the "Knit" button in RStudio to knit this file to a pdf.**

--------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, 
                      warning = FALSE, message = FALSE)
# add packages here 
library(readr)
library(tidyverse)
library(babynames)
library(stringr)
library(ggplot2)
```



## Problem 1: Regular expression 

### a. 

*answer:* There were 16314 baby names that end in vowels in 2017

```{r}
pattern <- "(.{1,})[aeiouy]\\b"
babynames %>%
  filter(year == 2017) %>%
  select(name) %>%
  unlist() -> names
sum(str_count(names, pattern))

```

### b. 

*answer:* 4 names matched this pattern, Edward, Eduard, Edgard, Edvard

```{r}
pattern2 <- "^Ed(.{1,})rd$"
x <- c("Edward", "Eddard", "Ned")
str_view(x, pattern2)
str_view(names, pattern2)
sum(str_count(names, pattern2))
```


### c. 

*answer:* Yes, there are 19 names like that, the names are Aya, Aiya, Yui, Iya, Aoi, Yue, Yoyo, Yu, Yiyi, Io, Ai, Yaa, Yi, Aaya, Oya, Yi, Yao, Yu, Ayo

```{r}
pattern3 <- "^([aeiouyAEIOUY]{1,})$"
str_view(names, pattern3)
sum(str_count(names, pattern3))
```

### d. 

*answer:* The length of the vector is 2960

```{r}
pattern4 <- "^.{4}$"
result <- babynames %>%
  filter(year == 2017) %>%
  mutate(is_4_rows = str_detect(name, pattern4)) %>% 
  filter(is_4_rows) %>% 
  dplyr::pull(name)

result %>% 
  length()

```



### e. 

*answer:* The palindrome names are anna, elle, emme, adda, alla, izzi, luul, avva, otto, anna

```{r}
pattern5 <- "(.)(.)\\2\\1"
result %>%
  str_to_lower() %>%
  str_subset(pattern5)
```


------------------------------------------------

## Problem 2: Energy autocorrelation

### a. 

```{r}
energy <- readr::read_csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/energy.csv",
                    col_type = cols(
                     .default = col_double(), 
                      Timestamp = col_datetime(format = ""),
                      dayWeek = col_factor(levels=c("Mon","Tues","Wed","Thurs","Fri","Sat","Sun"))
                     ))

```



```{r}
x <- energy %>% 
  arrange(Timestamp) %>%    # making sure sorted by time
  pull("Olin_Hall_of_Science")
acf_out <- acf(
  x,   # time series
  na.action = na.pass,    # skips over NAs
  lag.max = 4,   # max lag
  plot = FALSE)   # don't plot
acf_out
acf_out$acf   # autocorr values
acf_out$lag   # lag values
```

### a. 


*answer:* Function written

```{r}
autocor_fun <- function(KWHVect, maxLag){
  acf_out <- acf(
    KWHVect, #Time series
    na.action = na.pass,
    lag.max = maxLag,
    plot = FALSE
    )
  
  
  return(data.frame(autocor = as.vector(acf_out$acf), 
                    lag = 0:maxLag))
}
autocor_fun(x, 4)
```

### b. 

*answer:* write your answer here

```{r}
energy %>% 
  arrange(Timestamp) %>%    # making sure sorted by time
  select("Sayles-Hill" ,"Language_&_Dining_Center", "Olin_Hall_of_Science") %>% 
  map_df(autocor_fun, maxLag = 4, .id = "buildings")

```

### c. 

*answer:* write your answer here

```{r}
energy_narrow <- energy %>% 
  mutate(month = month(month, label=TRUE)) %>%
  pivot_longer(
    cols = 9:90,
    names_to = "building", 
    values_to = "energyKWH") 

#Special thanks to Jenna from the Stats Help Lab
energy_narrow %>% 
  group_by(building) %>% 
  arrange(building, Timestamp) %>% 
  filter(building == "Sayles-Hill" | 
           building == "Language_&_Dining_Center" | 
           building == "Olin_Hall_of_Science") %>% 
  reframe(auto_correlation = autocor_fun(energyKWH, 4)) %>% 
  unnest(auto_correlation)

```


### d. 

*answer:* When the lag is 0, the auto correlation values are very high, over 0.80, then they start to decrease in a shape that resembles a parabola. 

The auto correlation values reach a minimum at around 50, where they are either close to 0, or below 0. 

After that, the auto correlation value starts to increase, in a shape similar to a parabola, and gets to values close to 0.8.


```{r}
energy %>% 
  arrange(Timestamp) %>%    # making sure sorted by time
  select("Sayles-Hill" ,
         "Language_&_Dining_Center",
         "Olin_Hall_of_Science") %>% 
  map_df(autocor_fun,
         maxLag = 96,
         .id = "buildings") %>% 
  ggplot(aes(x = lag, y = autocor, color = buildings)) +
  geom_line() + labs(y = "auto correlation values", x = "lag")
  
```

------------------------------------------------


## Problem 3: weather

### a. 

*answer:* 

```{r}
library(nasaweather)
for(coeff in 5:11){
  str_c(colnames(atmos)[coeff], 
        (sd(atmos[[coeff]],
             na.rm = TRUE)/
        mean(atmos[[coeff]],
             na.rm = TRUE)), sep = " ") ->output
  print(output)
}
```

### b. 

*answer:* write your answer here

```{r}
coefficient_of_variation <- function(x){
  sd(x, na.rm = TRUE)/mean(x, na.rm = TRUE)
}
atmos %>% 
  select(-(1:4)) %>% 
  map_dfc(coefficient_of_variation)
```


### c. 


```{r}
# e.g. named vector with names x and y and values 1 and 2
c(x = 1, y = 2)
```


*answer:* 

```{r}
my_stats <- function(vect){
  c(mean = mean(vect, na.rm = TRUE),
    sd = sd(vect, na.rm = TRUE),
    min = min(vect, na.rm = TRUE),
    Q1 = quantile(vect, 0.25, na.rm = TRUE),
    median = median(vect, na.rm = TRUE),
    Q3 = quantile(vect, 0.75, na.rm = TRUE),
    max = max(vect, na.rm = TRUE))  
}
my_stats(c(1,2,3,4,5))
```

### d. 

*answer:* write your answer here

```{r}
atmos %>% 
  select(-(1:4)) %>% 
  map_df(my_stats) %>%
  mutate(variables = (colnames(atmos)[-(1:4)]))
```



### e. 

*answer:* 

```{r}
#I worked with Sarah and Allison for this question
atmos %>% 
  group_by(year) %>% 
  summarise(my_stats = my_stats(temp)) %>% 
  mutate(stat = c("mean", "sd", "min", "Q1", "median","Q3", "max")) %>% 
  pivot_wider(names_from = "stat",
              values_from = "my_stats")
```



------------------------------------------------------------------------

## Problem 4: 

```{r}
#install.packages("titanic")
library(titanic)
set.seed(12233)
df = tibble(titanic_train)  #load dataset
```

### a. 

*answer:*

```{r}
reg <- "(?<=\\s)[:alnum:]+(?=\\.)"
str_extract(df$Name, reg) ->titanicTitles
```


### b. 

*answer:*

```{r}
titanic_titles_df <- data.frame(titles = titanicTitles)
titanic_titles_df %>% 
  mutate(titles = factor(titles)) %>% 
  ggplot(aes(x = fct_infreq(titles))) +
  geom_bar(fill = "cornflowerblue") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 
#Text rotated based on code from https://forum.posit.co/t/boxplot-how-to-rotate-x-axis-labels-to-45/87126/3
```

### c.  

*answer:*

```{r}
reg <- ".+(?=\\,)"
str_extract(df$Name, reg) -> titanicLastNames
last_names_df <- data.frame(last_name = titanicLastNames)
most_popular_last_names <- last_names_df %>% 
  group_by(last_name) %>% 
  summarise(popular_last_name = n()) %>% 
  arrange(desc(popular_last_name)) %>% 
  slice(1:10)
most_popular_last_names


last_names_df %>% 
  filter(last_name %in% most_popular_last_names$last_name) %>% 
  mutate(last_name = factor(last_name)) %>% 
  ggplot(aes(x = fct_infreq(last_name))) +
  geom_bar(fill = "yellow") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 
```


### d. 

*answer:*

```{r}
mostPopularByLetter <- function(df, firstLetter, lastLetter){
  reg <- ".+(?=\\,)"
  str_extract(df$Name, reg) -> allLastNames
  filterLetters <- str_c("^",
                          str_to_upper(firstLetter),
                          ".*",
                          str_to_lower(lastLetter),
                          "$")
  filteredNames <- str_extract(allLastNames, filterLetters)
  names_df <- data.frame(last_name = filteredNames) %>% 
    drop_na() %>% 
    mutate(last_name = factor(last_name)) %>% 
    ggplot(aes(x = fct_infreq(last_name))) + 
    geom_bar(fill = "pink") +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 
    
    
}
res <- mostPopularByLetter(df, "S", "m")
res
```



