---
title: "Homework 6"
output: pdf_document
---

## Name: Arthur Viegas Eguia

## I worked with: Jacob Aronson (from the stats help lab)

**Click the "Knit" button in RStudio to knit this file to a pdf.**

--------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, prompt=TRUE,comment=NULL,message=FALSE, include=TRUE, warning = FALSE)
library(tidyverse)
library(ggplot2)
library(readr)
library(ggthemes)
theme_set(theme_stata(base_size = 10))  # for nice looking plots
library(rvest)
library(tidytext)
library(tidyr)
library(purrr)
library(polite)
library(tidytext)
library(plotly)
library(palmerpenguins)
library(dplyr)
library(shiny)
```


## Problem 1: Crimes

Scrape the table of data found at https://en.wikipedia.org/wiki/List_of_United_States_cities_by_crime_rate and create a plot showing property crime rate (total property crime) vs. violent crime rate (total violent crime).  Identify outlier cities by using a plotting command similar to the one below. (Don't blindly use this without thinking about the column names.)

```{r}
crimes_table <- bow("https://en.wikipedia.org/wiki/List_of_United_States_cities_by_crime_rate", user_agent = "data enthusiast") %>% 
  scrape() %>%
  html_table() %>% 
  .[[1]]


crimes_table_clean <- crimes_table %>% 
  set_names(.[1,]) %>% 
  janitor::clean_names() %>% # Clean names
  slice(c(-1, -2)) %>% # Remove the first row
  mutate(across(everything(), ~na_if(.x, ""))) %>% # Convert empty strings to NA
  type.convert(as.is = TRUE) %>% # Convert columns to their most appropriate type
  mutate_at(vars(c(9, 14)), parse_number)

crimes <- data.frame(city = crimes_table_clean[[2]],
                     violent_crime = crimes_table_clean[[9]],
                     property_crime = crimes_table_clean[[14]])
```

```{r, eval = FALSE}
ggplot(crimes, aes(x = violent_crime, y = property_crime, label = city)) +
    geom_point() +
    geom_text(
      data = filter(crimes, violent_crime > 1500 | property_crime > 6500), 
      check_overlap = TRUE, size = 2.5, nudge_y = 40
    )
```



------------------------------------------------



## Problem 2: Movie scraping

```{r}
url <- "https://www.boxofficemojo.com/chart/ww_top_lifetime_gross/?offset=0&area=XWW"
```


### a. 

*answer:* There is a single table

```{r}
movies_df <- read_html(url) %>% 
  html_table() #The first item in the list is a tibble, so it is technically a data frame
class(movies_df[[1]])
```

### b. 

*answer:* write your answer here

```{r}
movies_df_clean <- movies_df[[1]] %>% 
  janitor::clean_names() %>% 
  rename(world_dollars = worldwide_lifetime_gross,
         domestic_dollars = domestic_lifetime_gross,
         domestic_percentage = domestic_percent,
         overseas_dollars = foreign_lifetime_gross, 
         overseas_percentage = foreign_percent) %>% 
  slice(1:100)
  
```


### c. 


*answer:* 

```{r}
movie_df_right_types <- movies_df_clean %>% 
  mutate(world_dollars = parse_number(world_dollars),
         domestic_dollars = parse_number(domestic_dollars),
         domestic_percentage = parse_number(domestic_percentage),
         overseas_dollars = parse_number(overseas_dollars),
         overseas_percentage = parse_number(overseas_percentage))
movie_df_right_types
```




### d. 

*answer:* The titanic position is going to be 26, and the URL is https://www.boxofficemojo.com/title/tt0120338/?ref_=bo_cso_table_4

```{r}
titanic_position <- read_html(url) %>% 
  html_nodes("a") %>% 
  html_text() %>% 
  str_to_lower() %>% 
  str_which("titanic")
  
```

```{r}
titanic_href <- str_c("https://www.boxofficemojo.com",
                      read_html(url) %>% 
                      html_nodes("a") %>% 
                        html_attr("href")%>%
                        .[titanic_position])
titanic_href
```


### e. 

```{r}
temp_url <- "https://www.boxofficemojo.com/chart/ww_top_lifetime_gross/?offset=#&area=XWW"

```


*answer:* Indeed,"https://www.boxofficemojo.com/chart/ww_top_lifetime_gross/?offset=200&area=XWW" (which is just adding offset = 200) takes me to the page with movies starting at 201. 

```{r}
pull_table <- function(url_input){
  modified_df <- read_html(url_input) %>% 
  html_table() %>%
    .[[1]] %>% 
    janitor::clean_names() %>% 
    rename(world_dollars = worldwide_lifetime_gross,
          domestic_dollars = domestic_lifetime_gross,
          domestic_percentage = domestic_percent,
          overseas_dollars = foreign_lifetime_gross, 
          overseas_percentage = foreign_percent) %>% 
          mutate(across(c(3, 4, 5, 6, 7), parse_number))
  ranks <- c()
  for(i in modified_df$rank){
    if(is.character(i)){
      ranks <- c(ranks, parse_number(i))
    }
    else{
      ranks <- c(ranks, as.numeric(i))
    }
  }
  modified_df$rank <- ranks
  return(modified_df)
}

test_url <- "https://www.boxofficemojo.com/chart/ww_top_lifetime_gross/?offset=800#&area=XWW"

pull_table(test_url)
```



### f. 

*answer:* 

```{r}
all_url <- "https://www.boxofficemojo.com/chart/ww_top_lifetime_gross/?offset="
url_last_part <- "#&area=XWW"
idx <- seq(0, 800, by = 200)
movie_df <- map_df(idx, ~{
  new_url <- str_glue("{all_url}{.x}{url_last_part}")
  pull_table(new_url)
})
glimpse(movie_df)
head(movie_df)
```


## Problem 3: Penguins

```{r}
ui <- fluidPage(
  plotOutput("plot", height = 500)
)

server <- function(input, output){
  output$plot <- renderPlot({
    g <- ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g)) 
    g + geom_point()
  })
}

# you can modify the height to avoid scrolling
shinyApp(ui, server, options = list(height = 600))
```


### a. 

*answer:*

```{r}
ui <- fluidPage(
  checkboxInput("check", "Include color?", value = FALSE),
  plotOutput("plot", height = 500)
)

server <- function(input, output){
  output$plot <- renderPlot({
    if(input$check){
      g <- ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g, color = species)) 
    }
    else{
       g <- ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g))
    }
    g + geom_point()
  })
}

# you can modify the height to avoid scrolling
shinyApp(ui, server, options = list(height = 600))
```


### b. 

*answer:*

```{r}
ui <- fluidPage(
  checkboxInput("check", "Include color?", value = FALSE),
  varSelectInput("selectInputX", "X value", selected = "bill_length_mm", data = penguins %>% select(3:6)),
  varSelectInput("selectInputY", "Y value", selected = "body_mass_g", data = penguins %>% select(3:6)),
  plotOutput("plot", height = 500)
)

server <- function(input, output){
  output$plot <- renderPlot({
    if(input$check){
      g <- ggplot(penguins, aes(x = !!input$selectInputX, y = !!input$selectInputY, color = species)) 
    }
    else{
       g <- ggplot(penguins, aes(x = !!input$selectInputX, y = !!input$selectInputY))
    }
    g + geom_point()
  })
}

# you can modify the height to avoid scrolling
shinyApp(ui, server, options = list(height = 600))
```



### c. 


*answer:*

```{r}
ui <- fluidPage(
  checkboxInput("check", "Include color?", value = FALSE),
  varSelectInput("selectInputX", "X value", selected = "bill_length_mm", data = penguins %>% select(3:6)),
  varSelectInput("selectInputY", "Y value", selected = "body_mass_g", data = penguins %>% select(3:6)),
  plotOutput("plot", height = 500, click = "my_click"),
  dataTableOutput("my_table")
)

server <- function(input, output){
  output$plot <- renderPlot({
    if(input$check){
      g <- ggplot(penguins, aes(x = !!input$selectInputX, y = !!input$selectInputY, color = species)) 
    }
    else{
       g <- ggplot(penguins, aes(x = !!input$selectInputX, y = !!input$selectInputY))
    }
    g + geom_point()
  })
  output$my_table <- renderDataTable({
    nearPoints(penguins, input$my_click)
  })
}

# you can modify the height to avoid scrolling
shinyApp(ui, server, options = list(height = 600))
```




### d. 

*answer:*

```{r}
ui <- fluidPage(
  checkboxInput("check", "Include color?", value = FALSE),
  varSelectInput("selectInputX", "X value", selected = "bill_length_mm", data = penguins %>% select(3:6)),
  varSelectInput("selectInputY", "Y value", selected = "body_mass_g", data = penguins %>% select(3:6)),
  plotOutput("plot", height = 500, brush = "my_brush"),
  dataTableOutput("my_table")
)

server <- function(input, output){
  output$plot <- renderPlot({
    if(input$check){
      g <- ggplot(penguins, aes(x = !!input$selectInputX, y = !!input$selectInputY, color = species)) 
    }
    else{
       g <- ggplot(penguins, aes(x = !!input$selectInputX, y = !!input$selectInputY))
    }
    g + geom_point()
  })
  output$my_table <- renderDataTable({
    brushedPoints(penguins, input$my_brush)
  })
}

# you can modify the height to avoid scrolling
shinyApp(ui, server, options = list(height = 600))
```


------------------------------------------------



## Problem 4: Storm paths


Special thanks to Jacob Aaronson

```{r}
ctry <- map_data("world", 
                 region = c(
                   "usa",
                   "mexico",
                   "canada",
                   "uk"
                 ))
base_map <- ggplot(ctry) +  
  geom_polygon(aes(x = long,  y = lat, group = group)) + 
  labs( 
    x = "longitude", 
    y = "latitude", 
    title = "Atlantic storms paths"
    )  

base_map + 
  geom_path(data = storms, aes(x = long, y = lat, group = name), color = "red") +  
  coord_map(xlim  = c(min(storms$long), max(storms$long)),
            ylim  = c(min(storms$lat), max(storms$lat)))
head(storms)

```

### a.


*answer:*


```{r}
Individualstorm <- storms %>% 
      filter(name == "Amy") %>% 
  group_by(name) %>% 
  mutate(date = make_datetime(year = year,
                              month = month,
                              day = day,
                              hour = hour)) %>% 
  mutate(ellapsed_time = difftime(date, min(date), units = "hours")) %>% 
  mutate(ellapsed_time = as.numeric(ellapsed_time))

ui <- fluidPage(
  sliderInput("mySlider", "Storm Time", min = 0, max = 180, value = 90, step = 6),
  plotOutput("plot", height = 500)
  
)

server <- function(input, output){
  
  ellapsedTime <- reactive({filter(Individualstorm, ellapsed_time <= input$mySlider)})
  output$plot <- renderPlot({
    base_map +  geom_path(data = ellapsedTime(), aes(x = long, y = lat, group = name), color = "red")
  })
}

# you can modify the height to avoid scrolling
shinyApp(ui, server, options = list(height = 600))
```


### b. 

*answer:*

```{r}
ui <- fluidPage(
  selectInput("selectStorm", "Storm Name", selected = "Amy", choices = unique(storms$name)),
  plotOutput("plot", height = 500)
)

server <- function(input, output){
  
  stormOfInterest <- reactive({filter(storms, name == input$selectStorm)})
  output$plot <- renderPlot({
    base_map + geom_path(data = stormOfInterest(), aes(x = long, y = lat), color = "red")
  })
}

# you can modify the height to avoid scrolling
shinyApp(ui, server, options = list(height = 600))
#  coord_map(xlim  = c(min(long), max(long)),
#            ylim  = c(min(lat), max(lat)), data = stormOfInterest)

```


